---
title: "Interactive Visualization - Display time series plot of the COVID cases for each county in California"
author: "NINH NGUYEN"
date: "12/15/2020"
output: html_document
---

```{r setup, include=FALSE}
library(htmltools)
library(htmlwidgets)
library(maps)
library(leaflet)
library(XML)
library(xml2)
library(RCurl)
library(tidyverse)
library(rvest)
library(sf)
library(tigris)
library(gganimate)
library(transformr)  # to use polygons in animation
library(ggplot2)
```
CA covid data
```{r}
covid <- readRDS("~/ca.counties.covid.rds")
```

## Interact plot
### Draw a time series plot of the COVID cases for each county in California.

```{r}
library(plotly)
fig <- plot_ly(
  type = "scatter",
  x = as.Date(covid$date, format= "%Y-%m-%d"), 
  y = covid$cases,
  name = covid$name,
  mode = "markers",
)
fig <- fig %>%
  layout(
    title = "Time series plot of the COVID cases for each county in California every day from 2020-03-01 to 2020-12-10",
    xaxis = list(
      type = "date",
      range=c('2020-03-01', '2020-12-10')
      )
    )
saveWidget(fig, "Plot1 - Covid cases",selfcontained = FALSE)

```
• Have the tooltips for each observation show the date, number of cases and deaths for that day and the pro-rated values for cases and deaths relative to the county’s population, and the cumulative number of cases and deaths up to that date.
• Have the tooltips for each observation show the date, number of cases and deaths for that day and the pro-rated values for cases and deaths relative to the county’s population, and the cumulative number of cases and deaths up to that date.
• Allow the viewer to highlight one county, with the others still appearing but in the background.
• Allow the viewer to select a few counties and have these appear, but none of the others.
• Allow the viewer to toggle the time series for all counties, and then select a few to add them back to the plot. (The point is to not require the user to hide each county individually.)

```{r}
covid <- covid %>%
  group_by(name) %>%
  arrange(date) %>%
  mutate(
    cm_deaths = cumsum(deaths),
    cm_cases = cumsum(cases)
  )
```

```{r}
covid$hover <- with(covid,paste(
  '<br>',covid$date,
  '<br>',"Deaths:",covid$deaths,
  '<br>',"Cumulative Deaths:",cm_deaths,
  "<br>","pro-rated deaths per 10000:",covid$deaths*10000/covid$population,
  '<br>',"Cases:",covid$cases,
  '<br>',"Cumulative Cases:",cm_cases,
  "<br>","pro-rated cases per 10000:",covid$cases*10000/covid$population
                                )
  )
fig <- plot_ly(
  type = "scatter",
  x = as.Date(covid$date, format= "%Y-%m-%d"), 
  y = covid$cm_cases,
  name = covid$name,
  mode = "markers",
  hovertemplate = covid$hover
)

fig <- fig %>%
  layout(
    title = "Time series plot of the COVID cases for each county in California every day from 2020-03-01 to 2020-12-10",
    xaxis = list(
      type = "date",
      range=c('2020-03-01', '2020-12-10')
      )
    )

h = HTML("<h2>California County - Cumulative Covid Cases and number of deaths </h2><p>
This plot shows the COVID cases for each county in California by date.
<li>Double click on a county name in the legend to hide all of the lines</li>
<li>Click on some county name in the legend to show the times series</li>
<li>Hover/mouse over a point to see the tooltip and more details about that day's data for that county</li> ")

py2 = prependContent(fig,h)
saveWidget(py2, "Cumulative Covid cases",selfcontained = FALSE)
```


## Interactive on Map
• Draw a map of the counties in CA.
• Color each county by the percent
• Animate the color for each county over time.

## Import CA counties shape file
```{r}
ca_county<- read_sf("shape-files/CA_Counties/CA_Counties_TIGER2016.shp") %>%
  st_transform("+proj=longlat +datum=WGS84")
```
Group by month, counties
```{r}
covid$month <- format(covid$date, "%m")
covidMonth <- covid %>% 
  group_by(month,name) %>%
  summarise(deaths = sum(deaths),
            cases = sum(cases))
covidDF <- inner_join(ca_county,covidMonth,by = c("NAME"="name"))
```

```{r}
bins <- c(-1, 42, 352, 2000,6000, 90000, Inf)
pal <- colorBin("YlOrRd", domain = covidDF$cases, bins = bins)

labels <- sprintf(
  "<strong>%s</strong><br/>%g people / mi<sup>2</sup>",
  covidDF$NAME, covidDF$cases
) %>% lapply(htmltools::HTML)



mapCase <- leaflet(covidDF) %>%
  setView(-96, 37.8, 4) %>%
  addPolygons(
    data = covidDF,
    fillColor = ~pal(cases),
    weight = 2,
    opacity = 1,
    color = "white",
    dashArray = "3",
    fillOpacity = 0.7,
    highlight = highlightOptions(
      weight = 5,
      color = "#666",
      dashArray = "",
      fillOpacity = 0.7,
      bringToFront = TRUE),
    label = labels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "15px",
      direction = "auto")) %>%
  addLegend(pal = pal, values = ~cases, opacity = 0.7, title = NULL,
    position = "topright") 

h = HTML("<h2> Maps of California County - Total Covid Cases</h2><p>
<li>Hover/mouse over a county on map to see total Covid Cases</li> ")

py3 = prependContent(mapCase,h)
saveWidget(py3, "Map of total Covid cases",selfcontained = FALSE)
```

Animate map over time
```{r}
library(ggmap)
library(maps)
library(gganimate)
library(gifski)
library(transformr)  # to use polygons in animation

```

```{r}
md <- map_data('county', 'California')
state.md <- map_data('state', 'California')
covid$month <- format(covid$date, "%m")
covidMonth <- covid %>% 
  group_by(month,name,population) %>%
  summarise(deaths = sum(deaths),
            cases = sum(cases))
covidMonth$subregion <- tolower(covidMonth$name)
covidMonth$month <- as.numeric(covidMonth$month)
covidMonth$Percent <- covidMonth$cases*10000/covidMonth$population
plotData <- merge(md,covidMonth)
plotData <- plotData %>% mutate(xloc =-78.5,yloc=18.75)


baseMap <- ggplot() +
  geom_polygon(data = plotData, aes(x=long, y = lat, fill=Percent, group = group), color="white") + 
  geom_polygon(data = state.md, aes(x=long, y=lat, group=group), color="black", fill=NA) +
  coord_fixed(1.3) +
  scale_fill_gradient(low = "white", high = "orange", na.value="grey80") +
  labs(title = "Map of the counties in CA by percent cases on 10000 people from 2020/03 to 2020/12")
  
animateMap <- baseMap +
  theme_void() +
  theme(legend.position=c(.92,.7), legend.text.align=0, plot.title=element_text(hjust=.1, face="bold"), plot.caption=element_text(hjust=0)) +
  labs(title = 'Month: {frame_time}') + 
  transition_time(month) +
  ease_aes('linear')
# save as gif
mapGIF <- animate(animateMap) 
anim_save("MPCmap.gif", animation=mapGIF)
```
https://gganimate.com/
